---
title: "Tutorial 1"
author: "Author: Matthew J. Crossley"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
        collapsed: false
        smooth_scroll: true
    toc_depth: 3
    fig_caption: yes
    # code_folding: show
    number_sections: false
    theme: cosmo
fontsize: 14pt
---

```{r setup, include=FALSE}
library(reticulate)
use_python('/Users/mq20185996/miniconda3/bin/python')
```

## Learning objectives
* Be able to do basic `python` programming.
* Create and use variables of basic data types (`float`, `int`, `str`, `bool`).
* Create and use basic container data types (`list`, `tuple`, `set`, `dict`).
* Create and understand `for` and `while` loops.
* Create and understand conditional flow using `if`, `elif`, `else`.
* Create and understand how to define and use your own custom functions.

## Practice problems
* It's a good idea to work through these on your own, but if
you get very stuck, solutions can be found
[here](tutorial_1_solutions.html)

```{python message=FALSE, echo=T}
import numpy as np
import matplotlib.pyplot as plt
```

* Create a `float` variable named `ans_1a` that contains
the value `3.14159`.

* Create an `int` variable named `ans_1b` that contains the
value `3`.

* Create a `bool` variable named `ans_1c` that contains the
value `True`.

* Create a `str` variable named `ans_1d` that contains the
value `hello world`.

* Create a `list` named `ans_1e` that contains the elements
`1`, `'a'`, and `True` in that order.

* If possible, append the value `10` to the `list` you
created in the previous problem and save the result in a
variable named `ans_1f`.If not possible, write `ans_1f =
'can't because lists are immutable'`. Be sure not to modify
the `ans_1e` variable while doing this.

* Create `tuple` named `ans_1g`  that contains the elements
`1`, `'a'`, and `True` in that order.

* If possible, append the value `10` to the `tuple` you
created in the previous problem and save the result in a
variable named `ans_1h`. If not possible, write `ans_1h =
'can't because tuples are immutable'`. Be sure not to modify
the `ans_1g` variable while doing this.

* Create a `dict` variable named `ans_1i` with one key named
`A` that contains the list `[1, 2, 3]` and another key named
`B` that contains the tuple `('a', 'b', 'c', 'd')`.

a. Consider the following code chunk:
```{python}
x = 5
for i in [1, 3, 5]:
  x = i + 2
```
* What is the value of `x` after the loop has finished executing?
* What is the value of `i` after the loop has finished executing?
* Modify the code chunk such that the final value of `x` is
  10 and `i` is 20.

b. Consider the following code chunk:
```{python, echo=T, eval=F}
x = 10
y = [1, 2, 3]
while x > 5:
  y.append(x)
  x = x - 1
```
* How many elements does `y` contain after the loop has finished?
* What is the value of the final element in `y` and can you figure
  this out without executing the code?
  
c. Consider the following code chunk:
```{r, echo=T, eval=F}
if x > y: 
  z = 10
else:
  z = 2
```
* What is the value of `z` if `x = 10` and `y = -10`
* What is the value of `z` if `x = -5` and `y = 5`
* What is the value of `z` if `x = 0` and `y = 0`

d. Consider the following code chunk:
```{python, echo=T, eval=F}
for i in range(10):
  x = i * 2
  if x > 5:
    break
```
* How many times will this loop run?
* What is the value of `i` and `x` when the loop stops?

e. consider the following code chunk:
```{python, echo=T, eval=F}
if x == y:
  z = 1
elif x > y:
  z = 2
elif x < y:
  z = 3
```
* What is the value of `z` if `x = 10` and `y = -10`
* What is the value of `z` if `x = -5` and `y = 5`
* What is the value of `z` if `x = 0` and `y = 0`

f. Consider the following code chunk:
```{python, echo=T, eval=F}
def f(x):
  y = x**2
  return y

z = f(3)
```
* What are the values of `x`, `y` and `z`?

g. Consider the following code chunk:
```{python, echo=T, evsl=F}
def f(x):
  y = x**2
  return y

def g(x):
  y = x - 2
  return y

z = f(g(4))
```
* What are the values of `x`, `y` and `z`?

h. Consider the following code chunk:
```{python, echo=T, eval=F}
def f(x, y, z):
  res = x + y - z
  return res
```
* Is `f(2, 1, -4)` a valid way to call the function `f`?
* Is `f(x=2, 1, -4)` a valid way to call the function `f`?
* What is the wisest way to call the function `f`?

## Work on problem set 1
* [Problem set 1](homework_1.html)